Bioverse: Potentially Observable Exoplanet Biosignature Patterns Under the UV Threshold Hypothesis for the Origin of Life

### 2025-03-25 src/tex/ms.tex
delete biosignature detection rate
compute summary stats for stars

% Author list
\author[0000-0001-8355-2107]{Martin Schlecker}
%\author[0000-0001-8355-2107]{Martin Schlecker} # issue with ORCID? https://github.com/showyourwork/showyourwork/issues/454
\affiliation{Steward Observatory, The University of Arizona, Tucson, AZ 85721, USA; \href{mailto:schlecker@arizona.edu}{schlecker@arizona.edu}}
\author[0000-0003-3714-5855]{D\'{a}niel Apai}
\affiliation{Steward Observatory, The University of Arizona, Tucson, AZ 85721, USA; \href{mailto:schlecker@arizona.edu}{schlecker@arizona.edu}}
\affiliation{Lunar and Planetary Laboratory, The University of Arizona, Tucson, AZ 85721, USA}
\author[0000-0003-3481-0952]{Antonin Affholder}
\affiliation{Department of Ecology and Evolutionary Biology, University of Arizona, Tucson AZ, USA}
\author[0000-0002-5147-9053]{Sukrit Ranjan}
\affiliation{Lunar and Planetary Laboratory, The University of Arizona, Tucson, AZ 85721, USA}
\affiliation{Blue Marble Space Institute of Science, Seattle, 98104, USA}
\author[0000-0002-5806-5566]{Régis Ferrière}
\affiliation{Department of Ecology and Evolutionary Biology, University of Arizona, Tucson AZ, USA}
\affiliation{Institut de Biologie de l'École Normale Supérieure, ENS, PSL, Paris, France}
\affiliation{International Research Laboratory for Interdisciplinary Global Environmental Studies (iGLOBES), CNRS, ENS, PSL, University of Arizona, Tucson AZ, USA}
\author[0000-0003-3702-0382]{Kevin K.\ Hardegree-Ullman}
\affiliation{Steward Observatory, The University of Arizona, Tucson, AZ 85721, USA; \href{mailto:schlecker@arizona.edu}{schlecker@arizona.edu}}
\affiliation{Caltech/IPAC-NASA Exoplanet Science Institute, 1200 E. California Blvd., MC 100-22, Pasadena, CA 91125, USA}
\author[0000-0002-3286-7683]{Tim Lichtenberg}
\affiliation{Kapteyn Astronomical Institute, University of Groningen, PO Box 800, 9700 AV Groningen, The Netherlands}
\author[0000-0003-3557-6256]{Stéphane Mazevet}
\affiliation{Observatoire de la Côte d'Azur, Université Côte d'Azur, Nice, France}

@@

\section{Introduction}
\label{sec:intro}
\rev{The probability of abiogenesis remains one of the fundamental questions in modern science.
As of now, it has been primarily explored through statistical arguments~\citep[e.g.,][]{Spiegel2012,Kipping2021,Lingam2024}.
At the same time, a} wide variety of scenarios for the origin of life have been proposed~\citep[e.g.,][]{Baross1985,Brasier2011,Mulkidjanian2012,Fox2013,Deamer2015,Westall2018}.
While we may still be far from conclusively testing these scenarios, new prospects in the search for conditions favorable to life have opened up by thinking of the origin of life as a planetary phenomenon and identifying global-scale environmental properties that might support pathways to life~\citep{Sasselov2020}.
In particular, specific planetary conditions are needed to create stockpiles of initial compounds for prebiotic chemistry; and planetary processes are required to trigger the prebiotic synthesis.
Such planetary conditions can be hypothesized for exoplanets located in the \gls{HZ} of their host star, with persistent liquid water on their surface.
For example, if deep-sea or sedimentary hydrothermalism is required for abiogenesis, then the insulation of an ocean from the planetary crust minerals (e.g., due to high-pressure ices) may reduce or eliminate the chances of life emerging~\citep[e.g.,][]{Baross1985}.
The alternate scenario of a surface locally subject to wet-dry cycles requires a planetary exposure to mid-range \gls{UV} irradiation, as a source of energy and an agent of selection in chemical evolution~\citep[e.g.,][]{Deamer2019}.
This “UV Threshold Hypothesis” states that \gls{UV} light in a specific wavelength range played a constructive role in getting life started on Earth~\citep{Ranjan2016,Ranjan2017c,Rimmer2018,Rapf2016}, and it could provide a probabilistic approach to the interpretation of possible future biosignature detections~\citep[e.g.,][]{Catling2018,Walker2018}.

@@

\subsubsection{Simulated star and planet sample}
We generated two sets of synthetic exoplanet populations, one for FGK-type stars and one for M-type stars.
The stellar samples are drawn from the Gaia Catalogue of Nearby Stars~\citep{Smart2021} with a maximum Gaia magnitude of \var{M_G_max} and a maximum stellar mass of \var{M_st_max} \SI{}{\Msun}.
We included stars out to a maximum distance $d_{\max}$ that depends on the required planet sample size.
Planets were generated and assigned to the synthetic stars following the occurrence rates and size/orbit distributions of \citet{Bergsten2022}.
Following \citet{Bixel2021}, we considered only transiting \glspl{EEC} with radii $0.8\, S^{0.25} < R < 1.4 $ that are within the \gls{HZ} (see Section~\ref{sec:met-hz}).
The lower limit was suggested as a minimum planet size to retain an atmosphere~\citep{Zahnle2017}.
For all survey simulations and hypothesis tests, we repeated the above in a Monte Carlo fashion to generate randomized ensembles of synthetic star and planet populations~\citep[][]{Bixel2021}.

@@

The \gls{HZ} describes a region around a star where a planet with Earth's atmospheric composition and climate feedbacks can maintain liquid water on its surface~\citep[e.g.,][]{Ramirez2017,Ramirez2018,MolLous2022,Spinelli2023,Tuchow2023}.
Here, we adopted orbital distance estimates that define the \gls{HZ} as the region between the runaway greenhouse transition, where the stellar instellation cannot anymore be balanced through infrared cooling to space~\citep{Ingersoll1969}, and the maximum greenhouse limit, corresponding to the maximum distance at which surface temperatures allowing liquid water can be maintained through a CO$_2$ greenhouse effect~\citep{Kasting1991,Kasting1993,Underwood2003,Kopparapu2013,Kopparapu2014}.
\rev{The exact boundaries of the habitable zone are known to be sensitive to the star’s luminosity, spectral type, the planet’s mass, and the planet’s atmospheric properties~\citep[e.g.,][]{Pierrehumbert2011,Ramirez2014,Ramirez2017,Ramirez2018b,Koll2019,Ramirez2018,Ramirez2020,Bonati2021,Chaverot2022,Turbet2023}.
Here, we adopted the commonly used parametrization in \citet{Kopparapu2014} to derive luminosity and planetary mass-dependent distance limits of the \gls{HZ} $a_\mathrm{inner}$ and $a_\mathrm{outer}$.
}

@@

So far, we have \rev{drawn random values from uniform distributions for} $F_\mathrm{NUV, min}$, and $F_\mathrm{NUV}$\rev{, and from a log-uniform distribution for $f_\mathrm{life}$}. % Actually it is more than just NUV, it is whatever index that integrates \gls{NUV} and HZ requiremetns into a single scalar variable for which a threshold exist
A high biosignature detection rate $f_\mathrm{life}$ increases the evidence (see Equation~\ref{eq:bayes_factor}) but a survey strategy cannot influence it.
The same is true for $F_\mathrm{NUV, min}$, where again higher values increase the evidence as the binomial distribution for $H_\mathrm{1}$ gets increasingly skewed and shifted away from the one for $H_\mathrm{null}$.
However, one might select exoplanets for which a biosignature test is performed based on \textit{a priori} available contextual information~\citep{catling2018exoplanet} in order to maximize the science yield of investing additional resources.
For instance, the distribution of $F_\mathrm{NUV}$ in the planet sample can be influenced by the survey strategy, and a targeted sampling approach could favor extreme values. % of $F_\mathrm{NUV}$ in the sample selection.
We model this by distributing $F_\mathrm{NUV}$ according to different $Beta$ functions and introduce a selectivity parameter $s\in]-1,1[$ such that $F_\mathrm{NUV} \sim Beta(1/10^s,1/10^s)$.
Figure~\ref{fig:semian_selectivity} shows how the probability of obtaining true strong evidence for $H_\mathrm{1}$ scales with selectivity $s$.
%Here, $s=0$ corresponds to a random uniform distribution.
%Compared to this case, a high selectivity ($s \sim 1$) can increase the probability of obtaining true strong evidence to $\gtrsim \SI{85}{\percent}$ for large samples.
For large samples, a high selectivity ($s \sim 1$) can increase the probability of obtaining true strong evidence from $\sim \SI{70}{\percent}$ for $s= 0$ (random uniform distribution) to $> \SI{90}{\percent}$.

@@

%Meaningful tests of the UV~Threshold Hypothesis may thus require a different survey strategy that increases the sample of characterized planets.
The \gls{HWO} is expected to characterize a sample of $\sim 25$ Earth analogs~\citep{Mamajek2023,Tuchow2024}.
Depending on the technical design, \life\ is expected to target 25–80 \glspl{EEC}~\citep{Kammerer2018,Quanz2022}, which could be just sufficient to constrain the UV~Threshold Hypothesis.

\rev{If the rate of life’s emergence and persistence $f_\mathrm{life}$ is at a \SI{1}{\percent} level or lower, no currently envisioned future exoplanet mission has projected sample sizes sufficient to test the UV~Threshold Hypothesis.
One possible} exception is the Nautilus Space Observatory concept~\citep{Apai2019,Apai2022}.
Nautilus aims to characterize up to $\sim 1000$ \gls{EEC} via transmission spectroscopy, building on an innovative optical technology.
To guide the definition of future biosignature surveys, it is important to refine predictions on the role of \gls{UV} radiation in prebiotic chemistry with both theoretical and experimental work.

@@

\rev{\citet{Ranjan2017b} speculated that UV photochemistry might be rate-limiting for abiogenesis, particularly on planets orbiting M dwarfs, due to their lower baseline UV fluxes.
This could delay abiogenesis by orders of magnitude, resulting in a continuous dependence of abiogenesis likelihood on UV flux.
However, recent studies challenge this view for the cyanosulfidic scenario.
\citet{Rimmer2021} calculated photochemical timescales on early Earth at 180--300 hours (7.5--12.5 days), significantly shorter than the timescale for stochastic geological events also required for the scenario~\citep{Rimmer2023}.
Even with 1000x slower photochemistry on M-dwarf planets, prebiotic photochemistry like sulfite photolysis would still occur on a geologically negligible timescale of 20--30 years, meaning that the photochemistry is unlikely to be rate limiting compared to stochastic geological processes, justifying the step-function model.
}

>>>>

% Author list
\author[0000-0001-8355-2107]{Martin Schlecker}
%\author[0000-0001-8355-2107]{Martin Schlecker} # issue with ORCID? https://github.com/showyourwork/showyourwork/issues/454
\affiliation{Steward Observatory, The University of Arizona, Tucson, AZ 85721, USA; \href{mailto:schlecker@arizona.edu}{schlecker@arizona.edu}}
\author[0000-0003-3714-5855]{D\'{a}niel Apai}
\affiliation{Steward Observatory, The University of Arizona, Tucson, AZ 85721, USA; \href{mailto:schlecker@arizona.edu}{schlecker@arizona.edu}}
\affiliation{Lunar and Planetary Laboratory, The University of Arizona, Tucson, AZ 85721, USA}
\author[0000-0003-3481-0952]{Antonin Affholder}
\affiliation{Department of Ecology and Evolutionary Biology, University of Arizona, Tucson AZ, USA}
\author[0000-0002-5147-9053]{Sukrit Ranjan}
\affiliation{Lunar and Planetary Laboratory, The University of Arizona, Tucson, AZ 85721, USA}
\affiliation{Blue Marble Space Institute of Science, Seattle, 98104, USA}
\author[0000-0002-5806-5566]{R\'{e}gis Ferri\`{e}re}
\affiliation{Department of Ecology and Evolutionary Biology, University of Arizona, Tucson AZ, USA}
\affiliation{Institut de Biologie de l'École Normale Supérieure, ENS, PSL, Paris, France}
\affiliation{International Research Laboratory for Interdisciplinary Global Environmental Studies (iGLOBES), CNRS, ENS, PSL, University of Arizona, Tucson AZ, USA}
\author[0000-0003-3702-0382]{Kevin K.\ Hardegree-Ullman}
\affiliation{Steward Observatory, The University of Arizona, Tucson, AZ 85721, USA; \href{mailto:schlecker@arizona.edu}{schlecker@arizona.edu}}
\affiliation{Caltech/IPAC-NASA Exoplanet Science Institute, 1200 E. California Blvd., MC 100-22, Pasadena, CA 91125, USA}
\author[0000-0002-3286-7683]{Tim Lichtenberg}
\affiliation{Kapteyn Astronomical Institute, University of Groningen, PO Box 800, 9700 AV Groningen, The Netherlands}
\author[0000-0003-3557-6256]{St\'{e}phane Mazevet}
\affiliation{Observatoire de la C\^{o}te d'Azur, Universit\'{e} C\^{o}te d'Azur, Nice, France}

@@

\section{Introduction}
\label{sec:intro}
\rev{The probability of abiogenesis remains one of the fundamental questions in modern science.
As of now, it has been primarily explored through statistical arguments~\citep[e.g.,][]{Spiegel2012,Kipping2021,Lingam2024}.
At the same time, a} wide variety of scenarios for the origin of life have been proposed~\citep[e.g.,][]{Baross1985,Brasier2011,Mulkidjanian2012,Fox2013,Deamer2015,Westall2018}.
While we may still be far from conclusively testing these scenarios, new prospects in the search for conditions favorable to life have opened up by thinking of the origin of life as a planetary phenomenon and identifying global-scale environmental properties that might support pathways to life~\citep{Sasselov2020}.
In particular, specific planetary conditions are needed to create stockpiles of initial compounds for prebiotic chemistry; and planetary processes are required to trigger the prebiotic synthesis.
Such planetary conditions can be hypothesized for exoplanets located in the \gls{HZ} of their host star, with persistent liquid water on their surface.
For example, if deep-sea or sedimentary hydrothermalism is required for abiogenesis, then the insulation of an ocean from the planetary crust minerals (e.g., due to high-pressure ices) may reduce or eliminate the chances of life emerging~\citep[e.g.,][]{Baross1985}.
The alternate scenario of a surface locally subject to wet-dry cycles requires a planetary exposure to mid-range \gls{UV} irradiation, as a source of energy and an agent of selection in chemical evolution~\citep[e.g.,][]{Deamer2019}.
This "UV Threshold Hypothesis" states that \gls{UV} light in a specific wavelength range played a constructive role in getting life started on Earth~\citep{Ranjan2016,Ranjan2017c,Rimmer2018,Rapf2016}, and it could provide a probabilistic approach to the interpretation of possible future biosignature detections~\citep[e.g.,][]{Catling2018,Walker2018}.

@@

\subsubsection{Simulated star and planet sample}
We generated two sets of synthetic exoplanet populations, one for FGK-type stars and one for M-type stars.
The stellar samples are drawn from the Gaia Catalogue of Nearby Stars~\citep{Smart2021} with a maximum Gaia magnitude of \var{M_G_max} and a maximum stellar mass of \var{M_st_max} \SI{}{\Msun}.
We included stars out to a maximum distance $d_{\max}$ that depends on the required planet sample size.
\revv{In our FGK sample, we find that K~dwarfs dominate with effective temperatures of $\var{median_T_eff_st_K_FGK}^{+\var{p84_T_eff_st_K_FGK}}_{-\var{p16_T_eff_st_K_FGK}}$~K and masses of $\var{median_M_st_K_FGK}^{+\var{p84_M_st_K_FGK}}_{-\var{p16_M_st_K_FGK}}$~$M_\odot$ (median and 16th/84th percentiles).
In the M~dwarf sample, the stars have effective temperatures of $\var{median_T_eff_st_M_M}^{+\var{p84_T_eff_st_M_M}}_{-\var{p16_T_eff_st_M_M}}$~K and masses of $\var{median_M_st_M_M}^{+\var{p84_M_st_M_M}}_{-\var{p16_M_st_M_M}}$~$M_\odot$.}

Planets were generated and assigned to the synthetic stars following the occurrence rates and size/orbit distributions of \citet{Bergsten2022}.
Following \citet{Bixel2021}, we considered only transiting \glspl{EEC} with radii $0.8\, S^{0.25} < R < 1.4 $ that are within the \gls{HZ} (see Section~\ref{sec:met-hz}).
The lower limit was suggested as a minimum planet size to retain an atmosphere~\citep{Zahnle2017}.
For all survey simulations and hypothesis tests, we repeated the above in a Monte Carlo fashion to generate randomized ensembles of synthetic star and planet populations~\citep[][]{Bixel2021}.

@@

The \gls{HZ} describes a region around a star where a planet with Earth's atmospheric composition and climate feedbacks can maintain liquid water on its surface~\citep[e.g.,][]{Ramirez2017,Ramirez2018,MolLous2022,Spinelli2023,Tuchow2023}.
Here, we adopted orbital distance estimates that define the \gls{HZ} as the region between the runaway greenhouse transition, where the stellar instellation cannot anymore be balanced through infrared cooling to space~\citep{Ingersoll1969}, and the maximum greenhouse limit, corresponding to the maximum distance at which surface temperatures allowing liquid water can be maintained through a CO$_2$ greenhouse effect~\citep{Kasting1991,Kasting1993,Underwood2003,Kopparapu2013,Kopparapu2014}.
\rev{The exact boundaries of the habitable zone are known to be sensitive to the star's luminosity, spectral type, the planet's mass, and the planet's atmospheric properties~\citep[e.g.,][]{Pierrehumbert2011,Ramirez2014,Ramirez2017,Ramirez2018b,Koll2019,Ramirez2018,Ramirez2020,Bonati2021,Chaverot2022,Turbet2023}.
Here, we adopted the commonly used parametrization in \citet{Kopparapu2014} to derive luminosity and planetary mass-dependent distance limits of the \gls{HZ} $a_\mathrm{inner}$ and $a_\mathrm{outer}$.
}

@@

So far, we have \rev{drawn random values from uniform distributions for} $F_\mathrm{NUV, min}$, and $F_\mathrm{NUV}$\rev{, and from a log-uniform distribution for $f_\mathrm{life}$}. % Actually it is more than just NUV, it is whatever index that integrates \gls{NUV} and HZ requiremetns into a single scalar variable for which a threshold exist
\revv{Higher $f_\mathrm{life}$ increase} the evidence (see Equation~\ref{eq:bayes_factor}) but a survey strategy cannot influence \revv{this parameter}.
The same is true for $F_\mathrm{NUV, min}$, where again higher values increase the evidence as the binomial distribution for $H_\mathrm{1}$ gets increasingly skewed and shifted away from the one for $H_\mathrm{null}$.
However, one might select exoplanets for which a biosignature test is performed based on \textit{a priori} available contextual information~\citep{catling2018exoplanet} in order to maximize the science yield of investing additional resources.
For instance, the distribution of $F_\mathrm{NUV}$ in the planet sample can be influenced by the survey strategy, and a targeted sampling approach could favor extreme values. % of $F_\mathrm{NUV}$ in the sample selection.
We model this by distributing $F_\mathrm{NUV}$ according to different $Beta$ functions and introduce a selectivity parameter $s\in]-1,1[$ such that $F_\mathrm{NUV} \sim Beta(1/10^s,1/10^s)$.
Figure~\ref{fig:semian_selectivity} shows how the probability of obtaining true strong evidence for $H_\mathrm{1}$ scales with selectivity $s$.
%Here, $s=0$ corresponds to a random uniform distribution.
%Compared to this case, a high selectivity ($s \sim 1$) can increase the probability of obtaining true strong evidence to $\gtrsim \SI{85}{\percent}$ for large samples.
For large samples, a high selectivity ($s \sim 1$) can increase the probability of obtaining true strong evidence from $\sim \SI{70}{\percent}$ for $s= 0$ (random uniform distribution) to $> \SI{90}{\percent}$.

@@

%Meaningful tests of the UV~Threshold Hypothesis may thus require a different survey strategy that increases the sample of characterized planets.
The \gls{HWO} is expected to characterize a sample of $\sim 25$ Earth analogs~\citep{Mamajek2023,Tuchow2024}.
Depending on the technical design, \life\ is expected to target 25--80 \glspl{EEC}~\citep{Kammerer2018,Quanz2022}, which could be just sufficient to constrain the UV~Threshold Hypothesis.

\rev{If the rate of life's emergence and persistence $f_\mathrm{life}$ is at a \SI{1}{\percent} level or lower, no currently envisioned future exoplanet mission has projected sample sizes sufficient to test the UV~Threshold Hypothesis.
One possible} exception is the Nautilus Space Observatory concept~\citep{Apai2019,Apai2022}.
Nautilus aims to characterize up to $\sim 1000$ \gls{EEC} via transmission spectroscopy, building on an innovative optical technology.
To guide the definition of future biosignature surveys, it is important to refine predictions on the role of \gls{UV} radiation in prebiotic chemistry with both theoretical and experimental work.

@@

\rev{\citet{Ranjan2017b} speculated that UV photochemistry might be rate-limiting for abiogenesis, particularly on planets orbiting M dwarfs, due to their low baseline UV fluxes.
This could delay abiogenesis by orders of magnitude, resulting in a continuous dependence of abiogenesis likelihood on UV flux.
However, recent studies challenge this view for the cyanosulfidic scenario.
\citet{Rimmer2021} calculated photochemical timescales on early Earth at 180--300 hours (7.5--12.5 days), significantly shorter than the timescale for stochastic geological events also required for the scenario~\citep{Rimmer2023}.
Even with 1000x slower photochemistry on M-dwarf planets, prebiotic photochemistry like sulfite photolysis would still occur on a geologically negligible timescale of 20--30 years, meaning that the photochemistry is unlikely to be rate limiting compared to stochastic geological processes, justifying the step-function model.
}
